Probabilistic Inference of Basic Stellar Parameters:\\ Application to Flickering Stars \altaffilmark{\titledag}

### 2015-08-26 version1.0/HBM.tex
adding equation

We model the stochastic relationship between $F_8$ and \logg\ or \rhostar,
accounting for the fact that there exists some intrinsic scatter in
the dependent variable, and including the heteroskedastic uncertainties on both
the dependent and independent variables.
There are two excellent reasons for modelling the relation stochastically;
firstly, if the intrinsic scatter is ignored and the relation between
variables is assumed to be deterministic, those data points with smaller
measurement uncertainties may have an unrepresentative greater weighting
during the fitting process.
Secondly, we are interested in producing probability distributions over stellar
densities and surface gravities, as opposed to point estimates, and propagating
these probability distributions through to subsequent analyses.
Several recent studies have required posterior Probability Distribution
Function (PDF) samples, in order to conduct their (hierarchical)
analyses \citep[e.g.][]{rogers:2015, foreman-mackey:2014, angus:2015}.

Including observational uncertainties on the independent variable within our
analysis is also important, because ordinary least squares regression methods
performed on data with two-dimensional uncertainties will result in a slope
that is biased towards zero
(e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
For a demonstration of the effects of neglecting intrinsic scatter and
two-dimensional uncertainties, see Kelly (2007).

The two models describing the relationship between $F_8$, \logg\ and \rhostar\
density can be written as
\begin{equation}
	\log_{10}(\rho_\star) \sim \mathcal{N}(\mu = \alpha + \beta
	\log_{10}(F_8), \sigma = \sigma_{\rho}),
\end{equation}
\label{eq:rho}
and
\begin{equation}
	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
	\sigma = \sigma_g)
\end{equation}
\label{eq:logg}
The free parameters of the two models are $\alpha$, $\beta$, $\sigma_{\rho}$,
$\gamma$, $\delta$, and $\sigma_g$.
These relationships can be described as Gaussian distributions with means
given by the equation of a straight line, and standard deviations which
describe the intrinsic scatter about the mean.

In order to account for the two-dimensional observational uncertainties in
this data set, we marginalize over the latent, `true' values of $F_8$,
\rhostar\ and \logg\ using the importance sampling method of
\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
likelhood, we expect any bias introduced by the importance sampling method to
be negligable.}.
In what follows we give an overview of this procedure, but encourage the reader
to refer back  to this original reference for a more detailed description of
the mathematics.
The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be thought
of as a single `draw' from an underlying (assumed Gaussian) probability
distribution, with mean equal to the `true' value\footnote{We use `true' to
mean the value that would have been observed, given infinitely high
signal-to-noise.} and standard deviation equal to the observational
uncertainty.
We compute the likelihood of the data given the model, marginalized over the
true values, using importance sampling.
We sample from the posterior PDFs of the true values of the variables,
conditioned on the observed values, and compute the likelihood of each of those
samples.
% These posterior PDFs could be produced by the astronomers who provide the
% catalog.
% If posterior samples are made available, they can be used in this stage of
% the inference.
% In the event that they are not provided however, they can be generated.
% For example, the $log(g)$ values used here are point estimates of the
% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
% {\it Kepler} light curves and stellar models.
We would use the posterior samples generated in previous model fitting process
if they were available, however generating new samples is an acceptable
approximation provided the posteriors are Gaussian (or log-Normal) and the
priors used by the previous fitters were uninformative.
Such an approximation {\it will not be necessary} for those studies using
\rhostar\ or \logg\ calculated using the newly-calibrated relations
presented here, as long as the posterior samples are published.
After producing new posterior PDF samples we then add up the individual
log-likelihoods of each sample to compute the total marginalized likelihood
for a star.
The marginalized likelihood of the whole dataset is then computed as the sum
of the log-likelihoods of each star.
The importance sampling method allows us to approximate the integral over the
latent variables.

Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
model.
The faint colored lines show the model where the parameters have been drawn
from their posterior PDFs.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
Stellar density vs. flicker. This figure shows the model, conditioned on the
data. The solid black line shows the median model values, and the pink lines
are random draws from the posterior PDFs of the model parameters; $\alpha$,
$\beta$ and $\sigma_{\rho}$.
}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
the model, conditioned on the data. The solid black line shows the median model
values, and the blue lines are random draws from the posterior PDFs of the
model parameters; $\gamma$, $\delta$ and $\sigma_g$.
}
\label{fig:logg}
\end{center}
\end{figure}

\begin{table}
\caption{Median parameter values with their 16th and 84th percentile
	 uncertainties.}
\begin{tabular}{lcc}
\hline\hline
    Parameter & Median value \\
    \hline
    $\alpha$         &   \\
    $\beta$          &   \\
    $\sigma_{\rho}$  &   \\
    $\gamma$         &   \\
    $\delta$         &   \\
    $\sigma_{g}$     &   \\
    \hline
\end{tabular}
\end{table}

>>>>

We model the stochastic relationship between $F_8$ and \logg\ or \rhostar,
accounting for the fact that there exists some intrinsic scatter in
the dependent variable, and including the heteroskedastic uncertainties on both
the dependent and independent variables.
There are two excellent reasons for modelling the relation stochastically;
firstly, if the intrinsic scatter is ignored and the relation between
variables is assumed to be deterministic, those data points with smaller
measurement uncertainties may have an unrepresentative greater weighting
during the fitting process.
Secondly, we are interested in producing probability distributions over stellar
densities and surface gravities, as opposed to point estimates, and propagating
these probability distributions through to subsequent analyses.
Several recent studies have required posterior Probability Density
Function (PDF) samples, in order to conduct their (hierarchical)
analyses \citep[e.g.][]{foreman-mackey:2014, rogers:2015, angus:2015}.

Including observational uncertainties on the independent variable within our
analysis is also important, because ordinary least squares regression methods
performed on data with two-dimensional uncertainties will result in a slope
that is biased towards zero
\citep[e.g.][]{fuller:1987, fox:1997}.
For a demonstration of the affects of neglecting intrinsic scatter and
two-dimensional uncertainties, see \citet{kelly:2007}.

We write the two models describing the relationship between $F_8$, \logg\ and
\rhostar\ as
\begin{equation}
	\log_{10}(\rho_\star) \sim \mathcal{N}(\mu = \alpha + \beta
	\log_{10}(F_8), \sigma = \sigma_{\rho}),
\end{equation}
\label{eq:rho}
and
\begin{equation}
	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
	\sigma = \sigma_g).
\end{equation}
\label{eq:logg}
The free parameters of the two models are $\alpha$, $\beta$, $\sigma_{\rho}$,
$\gamma$, $\delta$, and $\sigma_g$.
These relationships can be described as Gaussian distributions with means
given by the equation of a straight line, and standard deviations which
describe the intrinsic scatter about the mean.

In order to account for the two-dimensional observational uncertainties in
this data set, we marginalize over the latent, `true' values of $F_8$,
\rhostar\ and \logg\ using the importance sampling method of
\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
likelhood, we expect any bias introduced by the importance sampling method to
be negligable.}.
In what follows we give an overview of this procedure, but encourage the reader
to refer back  to this original reference for a more detailed description of
the mathematics.
The observed values of the variables; $F_8$, \rhostar\ and \logg\ can be thought
of as a single `draw' from an underlying (assumed Gaussian) probability
distribution, with mean equal to the `true' value\footnote{We use `true' to
mean the value that would have been observed, given infinitely high
signal-to-noise.} and standard deviation equal to the observational
uncertainty.
We compute the likelihood of the data given the model, marginalized over the
true values, using importance sampling.
We sample from the posterior PDFs of the true values of the variables,
conditioned on the observed values, and compute the likelihood of each of those
samples.
% These posterior PDFs could be produced by the astronomers who provide the
% catalog.
% If posterior samples are made available, they can be used in this stage of
% the inference.
% In the event that they are not provided however, they can be generated.
% For example, the $log(g)$ values used here are point estimates of the
% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
% {\it Kepler} light curves and stellar models.
We would use the posterior samples generated in previous model fitting process
if they were available, however generating new samples is an acceptable
approximation provided the posteriors are Gaussian (or log-Normal) and the
priors used by the previous fitters were uninformative.
Such an approximation {\it will not be necessary} for those studies using
\rhostar\ or \logg\ calculated using the newly-calibrated relations
presented here, since we have published our posterior samples.
After producing new posterior PDF samples we then add up the individual
log-likelihoods of each sample to compute the total marginalized likelihood
for a star.
The marginalized likelihood of the whole dataset is then computed as the sum
of the log-likelihoods of each star.
The importance sampling method allows us to approximate the integral over the
latent variables.
We used flat priors for all except the $\sigma$ parameters, for which we used
priors which were flat in log-space.

Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
models.
The shaded regions show the 1 and 2$\sigma$ confidence interval.
These intervals are representative of the uncertainty in the flicker relations,
but also of the intrinsic scatter in those relations.
% Given a measurement of flicker, the predictive probability distribution of
% \logg and \rhostar will be a Gaussian distribution

We also used an analytic form of the marginal likelihood, which accounts for
2-d uncertainties but does not allow for the modelling of the intrinsic scatter
as a function of $x$ or $y$.
For the relation between flicker and \rhostar This likelihood function can be
written as
\begin{eqnarray}
	p(\rho_*|\alpha, \beta, \sigma_{\rho}) = &
	\exp \left[-\frac{1}{2} \sum_{n=1}^N\frac{[y_n-(\alpha + \beta x_n)]^2}
	{\left[\beta \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2\right]}\right]
	\\ \nonumber
	& + \beta^2 \sigma_x^2 + (\sigma_y + \sigma_{\rho})^2,
\end{eqnarray}
and similarly for \logg.
The parameter values found using this simpler likelihood function are included
in table \ref{tab:results2}.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
Stellar density vs. flicker.
This figure shows the model, conditioned on the data.
The solid black line shows the highest-likelihood sample, and the pink shaded
regions show the 1 and 2$\sigma$ confidence intervals.}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
$\log(g)$ vs. flicker.
As in \ref{fig:rhostar} this figure shows the model, conditioned on the data.
The solid black line shows the highest-likelihood sample, and the blue shaded
regions show the 1 and 2$\sigma$ confidence intervals.}
\label{fig:logg}
\end{center}
\end{figure}

\begin{table}
\caption{Highest-likelihood parameter values with their 16th and 84th
	percentile uncertainties.}
\begin{tabular}{lcc}
\hline\hline
    Parameter & highest-likelihood value \\
    \hline
    $\alpha$         &   \\
    $\beta$          &   \\
    $\sigma_{\rho}$  &   \\
    $\gamma$         &   \\
    $\delta$         &   \\
    $\sigma_{g}$     &   \\
    \hline
\end{tabular}
\end{table}

### 2015-08-26 version1.0/manuscript.tex
adding equation

RA thanks...
DMK thanks...

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]{seager:2003}
Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}

>>>>

RA would like to thank Dan Foreman-Mackey for his extremely helpful suggestions
and comments, and Angie Wolfgang and John Johnson for their fantastic advice.
RA would also like to thank the SAMSI institute and the other members of the
SAMSI {\it Kepler} exoplanet statistics working groups.
DMK thanks...

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Fox}{1997}]{fox:1997}
Fox, J., 1997, Applied Regression Analysis, Linear Models, and Related Methods
(Thousand Oaks:Sage Publications, Inc.)
\bibitem[\protect\citeauthoryear{Fuller}{1987}]{fuller:1987}
Fuller, W.~A., 1987, Wiley Series in Probability and Mathematical Statistics
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kelly}{2007}]{kelly:2007} Kelly, B.~C., 2007,
ApJ, 665, 1489
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Seager \& Mall\'{e}n-Ornelas}{2003}]
{seager:2003}Seager, S., \& Mall\'{e}n-Ornelas, G., 2003, ApJ, 585, 1038
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}
